\Large {\bf Protocol-Guided Analysis of Post-silicon Traces Under Limited Observability}

### 2015-09-16 paper.tex
Starting with LPN

\author{\large Hao Zheng$^1$, Yuting Cao$^1$, Sandip Ray$^2$, Jin Yang$^2$ \\
$^1$Dept. of Computer Science and Eng., University of South Florida, Tampa, Fl 33620. USA. \\
$^2$Strategic CAD Labs, Intel Corporation, Hillsboro, OR 97124.  USA.}
\maketitle

@@

In this paper, we consider the problem of reconstructing
protocol-level behavior from silicon traces in SoC designs.
More specifically, given a collection of system-level
communication protocols and a trace of (partially observed)
hardware signals, our approach infers, with a certain
measure of confidence, the protocol instances (and their
interleavings) being exercised by the trace.  Our approach
is based on a formalization of system-level transactions via
labeled Petri-Nets, which are capable of
describing sequencing, concurrency, and choices of events.
Given this formalization, we develop algorithms to infer
system-level transactions from traces with missing, noisy,
and ambiguous signal values, together with an estimate of
confidence on the inference.  We demonstrate our approach on
two SoC models: a SystemC prototype and a more realistic
implementation constructed within the GEM5
environment~\cite{Binkert2011}.

@@

\begin{figure*}[tb]
\begin{center}
\begin{tabular}{cc}
\begin{minipage}{3in}
\includegraphics[height=2.5in]{figures/bpmn-flow-ex}
\end{minipage}
&
\begin{minipage}{3in}
\resizebox{2.8in}{!}{
\begin{tikzpicture}[node distance=3cm, auto,>=latex', thick]
\tikzset{rectbox/.style={rectangle, draw, align=flush center,thick, minimum size = 6mm}}
\tikzset{circ/.style={circle, draw,thick}}
\tikzset{terminal/.style={circle, draw,line width=.8mm}}
\tikzset{weight3/.style={line width=.3mm}}
	% Nodes:
	\node[circ]		(p1) 		at (0,0) {$p_1$};
	\node[rectbox]	(t1)	at	($(p1) + (0,-1.1)$) {$t_1:\langle {\tt Driver:Device:Load} \_fw\rangle$};
	\node[circ] 	(p2) 		at ($(t1) + (0,-1.1)$) {$p_2$};
	\node[rectbox]	(t2)	at	($(p2) + (0,-1.1)$) {$t_2:\langle {\tt Device:CE:Auth\_req} \rangle$};
	\node[circ] 	(p3) 		at ($(t2) + (0,-1.1)$) {$p_3$};
	\node[rectbox]	(t3)	at	($(p3) + (0,-1.1)$) {$t_3:\langle {\tt CE:Device:Auth\_resp}\rangle$};
	\node[circ] 	(p4) 		at ($(t3) + (0,-1.1)$) {$p_4$};
	\node[rectbox]	(t4)	at	($(p4) + (-3,-1.1)$) {$t_4:\langle {\tt Device:Driver:Report} \rangle$};
	\node[rectbox]	(t5)	at	($(p4) + (3,-1.1)$) {$t_5:\langle {\tt Device:CE:Ack} \rangle$};
	\node[circ,fill=black] 	(term1) 		at ($(t4) + (0,-1.1)$) {$~~~$};
	\node[circ,fill=black] 	(term2) 		at ($(t5) + (0,-1.1)$) {$~~~$};

@@

\end{tikzpicture}
}
\end{minipage}
\\
(a) & (b)
\end{tabular}
\caption{(a) A graphical representation of a SoC firmware load protocol in BPMN~\cite{Krstic14HOST}, and (b) the LPN capturing the communication protocol in (a).}
\label{flow-spec-ex}
\end{center}
\end{figure*}

\section{Background}

\subsection{System-level Transactions in SoC Designs}

An SoC design is typically composed of a large number of
pre-designed hardware or software blocks (often referred to
as ``intellectual properties'' or ``IPs'') that coordinate
through complex protocols to implement system-level
behavior.  Fig.~\ref{fig:flow-spec-ex} shows one such
protocol, involving th

\subsection{Labeled Petri-Nets}

Although system flows depicted in BPMN as in Fig.~\ref{flow-spec-ex} are intuitive for system architects to understand communications across multiple components, they are not suitable for algorithmic analysis.  In this section, we present an alternative formalism, \emph{labeled Petri Nets}, to represent communication aspects of system flows with the intuitive descriptions of internal operations for IP blocks abstracted away whenever possible.  Labeled Petri nets, which are a subclass of traditional Petri-nets~\cite{Murata89} where Petri-net transitions are labeled with user defined information,  are capable of describing sequencing, concurrency, and choices.   This formalism has well defined operational semantics, efficient analysis analysis algorithms and and has been used widely in modeling and analyzing communication protocols, concurrent programs, etc.

@@

Specifically, a labeled Petri-net (LPN) is a tuple $(P, T, s_0, E, L)$ where
\begin{itemize}
\item $P$ is a finite set of places,
\item $T$ is a finite set of transitions,
\item $\mathit{init}$ is the set of initially marked places, also referred to as the initial marking.
\item $E$ is a finite set of events,
\item $L: T \rightarrow E$ is a labeling function that maps each transition $t \in T$ to an event $e \in E$.
\end{itemize}

For each transition $t \in T$, its preset, denoted as $\bullet{t} \subseteq P$, is the set of places connected to $t$, and its postset, denoted as $t\bullet \subseteq P$, is the set of places that $t$ is connected to.  A marking $s \subseteq P$ of a LPN is a subset of places marked with tokens, and it is also referred to as a state of a LPN.  The initial marking $\mathit{init}$ is also the initial state of the LPN.

The operational semantics of a LPN is defined by transition executions.  A transition can be executed after it is {\em enabled}.  A transition $t \in T$ is enabled in a state $s$ if every place in its preset is included in the marking, i.e. $\bullet{t} \subseteq s$.  Execution of $t$ results in a new state $s'$ such that 
\[
s' = (s - \bullet{t}) \cup t\bullet,
\]
and the emission of event $e$ labeled for $t$.

The communication protocol in the BMPN shown in~Fig.~\ref{flow-spec-ex}(a) is represented by the LPN shown in~Figure~\ref{flow-spec-ex}(b).  In this and the following figures for LPNs, the labeled circles denote places, and the labeled boxes denote transitions.  Each transition is labeled with its name and the associated event.  Each event has a form of $\langle {\tt src, dest, cmd} \rangle$ where ${\tt cmd}$ is a command sent from a source component ${\tt src}$ to a destination component ${\tt dest}$. The solid black places without outgoing edges are {\em terminals}, which indicate termination of protocols represented by the LPNs.  The initial marking is $\mathit{init} = \{p_1\}$.  In this LPN model, only the communication portion of the BPMN specification is represented while the computation portion is ignored.

>>>>

%% \author{\large Hao Zheng$^1$, Yuting Cao$^1$, Sandip Ray$^2$, Jin Yang$^2$ \\
%% $^1$Dept. of Computer Science and Eng., University of South Florida, Tampa, Fl 33620. USA. \\
%% $^2$Strategic CAD Labs, Intel Corporation, Hillsboro, OR 97124.  USA.}
\maketitle

@@

In this paper, we consider the problem of reconstructing
protocol-level behavior from silicon traces in SoC designs.
More specifically, given a collection of system-level
communication protocols and a trace of (partially observed)
hardware signals, our approach infers, with a certain
measure of confidence, the protocol instances (and their
interleavings) being exercised by the trace.  Our approach
is based on a formalization of system-level transactions via
labeled Petri-Nets, which are capable of describing
sequencing, concurrency, and choices of events.  Given this
formalization, we develop algorithms to infer system-level
transactions from traces with missing, noisy, and ambiguous
signal values, together with an estimate of confidence on
the inference.  We demonstrate our approach on two SoC case
studies: a SystemC prototype and a more realistic
implementation constructed within the GEM5
environment~\cite{Binkert2011}.

@@

\begin{figure*}[tb]
\begin{center}
\begin{tabular}{cc}
\begin{minipage}{3in}
\includegraphics[height=2in]{figures/bpmn-flow-ex}
\end{minipage}
&
\begin{minipage}{3in}
\resizebox{2.8in}{!}{
\begin{tikzpicture}[node distance=3cm, auto,>=latex', thick]
\tikzset{rectbox/.style={rectangle, draw, align=flush center,thick, minimum size = 6mm}}
\tikzset{circ/.style={circle, draw,thick}}
\tikzset{terminal/.style={circle, draw,line width=.8mm}}
\tikzset{weight3/.style={line width=.3mm}}
	% Nodes:
	\node[circ]		(p1) 		at (0,0) {$p_1$};
	\node[rectbox]	(t1)	at	($(p1) + (0,-1.1)$) {$t_1:\langle {\tt Driver:Device:Load} \_fw\rangle$};
	\node[circ] 	(p2) 		at ($(t1) + (0,-1.1)$) {$p_2$};
	\node[rectbox]	(t2)	at	($(p2) + (0,-1.1)$) {$t_2:\langle {\tt Device:CE:Auth\_req} \rangle$};
	\node[circ] 	(p3) 		at ($(t2) + (0,-1.1)$) {$p_3$};
	\node[rectbox]	(t3)	at	($(p3) + (0,-1.1)$) {$t_3:\langle {\tt CE:Device:Auth\_resp}\rangle$};
	\node[circ] 	(p4) 		at ($(t3) + (0,-1.1)$) {$p_4$};
	\node[rectbox]	(t4)	at	($(p4) + (-3,-1.1)$) {$t_4:\langle {\tt Device:Driver:Report} \rangle$};
	\node[rectbox]	(t5)	at	($(p4) + (3,-1.1)$) {$t_5:\langle {\tt Device:CE:Ack} \rangle$};
	\node[circ,fill=black] 	(term1) 		at ($(t4) + (0,-1.1)$) {$~~~$};
	\node[circ,fill=black] 	(term2) 		at ($(t5) + (0,-1.1)$) {$~~~$};

@@

\end{tikzpicture}
}
\end{minipage}
\\
(a) & (b)
\end{tabular}
\caption{(a) A graphical representation of a SoC firmware
  load protocol~\cite{Krstic14HOST}.  (b) LPN formalization.
  Each event has a form of $\langle {\tt src, dest, cmd}
  \rangle$ where ${\tt cmd}$ is a command sent from a source
  component ${\tt src}$ to a destination component ${\tt
    dest}$. The solid black places without outgoing edges
  are {\em terminals}, which indicate termination of
  protocols represented by the LPNs.}
\label{flow-spec-ex}
\end{center}
\end{figure*}

\section{Labeled Petri-Nets}

Labeled Petri-nets (LPN) is a formalization of state
transition systems that is capable of describing sequencing,
concurrency, and choices.  Fig.~\ref{flow-spec-ex} shows the
formalization of a simple SoC protocol with Petri-Nets.
Formally, an LPN is a tuple $(P, T, s_0, E, L)$ where $P$ is
a finite set of {\em places}, $T$ is a finite set of {\em
  transitions}, $\mathit{init}$ is the set of initially
marked places, also referred to as the {\em initial
  marking}, $E$ is a finite set of {\em events}, and $L: T
\rightarrow E$ is a labeling function that maps each
transition $t \in T$ to an event $e \in E$.  For each
transition $t \in T$, its preset, denoted as $\bullet{t}
\subseteq P$, is the set of places connected to $t$, and its
postset, denoted as $t\bullet \subseteq P$, is the set of
places that $t$ is connected to.  A marking $s \subseteq P$
of a LPN is a subset of places marked with tokens, and it is
also referred to as a state of a LPN.  The initial marking
$\mathit{init}$ is also the initial state of the LPN.

%% This formalism has well defined
%% operational semantics, efficient analysis analysis
%% algorithms and and has been used widely in modeling and
%% analyzing communication protocols, concurrent programs, etc.

@@

%% Formally, an LPN is a tuple $(P, T, s_0, E, L)$ where $P$ is
%% a finite set of {\em places}, $T$ is a finite set of {\em
%%   transitions}, $\mathit{init}$ is the set of initially
%% marked places, also referred to as the {\em initial
%%   marking}, $E$ is a finite set of {\em events}, and $L: T
%% \rightarrow E$ is a labeling function that maps each
%% transition $t \in T$ to an event $e \in E$.  For each
%% transition $t \in T$, its preset, denoted as $\bullet{t}
%% \subseteq P$, is the set of places connected to $t$, and its
%% postset, denoted as $t\bullet \subseteq P$, is the set of
%% places that $t$ is connected to.  A marking $s \subseteq P$
%% of a LPN is a subset of places marked with tokens, and it is
%% also referred to as a state of a LPN.  The initial marking
%% $\mathit{init}$ is also the initial state of the LPN.

%% The operational semantics of a LPN is defined by transition executions.  A transition can be executed after it is {\em enabled}.  A transition $t \in T$ is enabled in a state $s$ if every place in its preset is included in the marking, i.e. $\bullet{t} \subseteq s$.  Execution of $t$ results in a new state $s'$ such that 
%% \[
%% s' = (s - \bullet{t}) \cup t\bullet,
%% \]
%% and the emission of event $e$ labeled for $t$.

%% The communication protocol in the BMPN shown
%% in~Fig.~\ref{flow-spec-ex}(a) is represented by the LPN
%% shown in~Figure~\ref{flow-spec-ex}(b).  In this and the
%% following figures for LPNs, the labeled circles denote
%% places, and the labeled boxes denote transitions.  Each
%% transition is labeled with its name and the associated
%% event.  Each event has a form of $\langle {\tt src, dest,
%%   cmd} \rangle$ where ${\tt cmd}$ is a command sent from a
%% source component ${\tt src}$ to a destination component
%% ${\tt dest}$. The solid black places without outgoing edges
%% are {\em terminals}, which indicate termination of protocols
%% represented by the LPNs.  The initial marking is
%% $\mathit{init} = \{p_1\}$.  In this LPN model, only the
%% communication portion of the BPMN specification is
%% represented while the computation portion is ignored.
